### 2016-03-03 OSS_2016/referenc.tex
change references format

%%%%%%%%%%%%%%%%%%%%%%%% referenc.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% sample references
% "computer science"
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%

%
% BibTeX users please use
% \bibliographystyle{}
% \bibliography{}
%
% Non-BibTeX users please use
\begin{thebibliography}{99.}
%
% and use \bibitem to create references.
%
% Use the following syntax and markup for your references
%
% Monographs
%\bibitem{monograph} Kajan E (2002)
%Information technology encyclopedia and acronyms. Springer, Berlin
%Heidelberg New York

% Contributed Works
%\bibitem{contribution} Broy M (2002) Software engineering -- From
%auxiliary to key technologies. In: Broy M, Denert E (eds)
%Software Pioneers. Springer, Berlin Heidelberg New York
\bibitem{Antoniol}Antoniol, G., Ayari, K., Di Penta, M., Khomh, F., \& Gu\'eh\'eneuc, Y. G. (2008, October). Is it a bug or an enhancement?: a text-based approach to classify change requests. In Proceedings of the 2008 conference of the center for advanced studies on collaborative research: meeting of minds (p. 23). ACM.
\bibitem{Herzig}Herzig, K., Just, S., \& Zeller, A. (2013, May). It's not a bug, it's a feature: how misclassification impacts bug prediction. In Proceedings of the 2013 International Conference on Software Engineering (pp. 392-401). IEEE Press.
\bibitem{Sliwerski}J. Śliwerski, J., Zimmermann, T., \& Zeller, A. (2005, May). When do changes induce fixes?. In ACM sigsoft software engineering notes (Vol. 30, No. 4, pp. 1-5). ACM.
\bibitem {Nguyen}Nguyen, A. T., Nguyen, T. T., Nguyen, H. A., \& Nguyen, T. N. (2012, November). Multi-layered approach for recovering links between bug reports and fixes. In Proceedings of the ACM SIGSOFT 20th International Symposium on the Foundations of Software Engineering (p. 63). ACM.
\bibitem {Zimmermann}Zimmermann, T., Premraj, R., \& Zeller, A. (2007, May). Predicting defects for eclipse. In Predictor Models in Software Engineering, 2007. PROMISE'07: ICSE Workshops 2007. International Workshop on (pp. 9-9). IEEE.
\bibitem{Thomas}Zimmermann, T., \& Weißgerber, P. (2004, May). Preprocessing CVS data for fine-grained analysis. In Proceedings of the First International Workshop on Mining Software Repositories (pp. 2-6). sn.
\bibitem{Bird}Bird, C., Bachmann, A., Aune, E., Duffy, J., Bernstein, A., Filkov, V., \& Devanbu, P. (2009, August). Fair and balanced?: bias in bug-fix datasets. In Proceedings of the the 7th joint meeting of the European software engineering conference and the ACM SIGSOFT symposium on The foundations of software engineering (pp. 121-130). ACM.
\bibitem{NguyenTH}Nguyen, T. H., Adams, B., \& Hassan, A. E. (2010, October). A case study of bias in bug-fix datasets. In Reverse Engineering (WCRE), 2010 17th Working Conference on (pp. 259-268). IEEE.
% Journal
%\bibitem{journal} Che M, Grellmann W, Seidler S (1997)
%Appl Polym Sci 64:1079--1090

% Theses
%\bibitem{thesis} Ross DW (1977) Lysosomes and storage diseases. MA
%Thesis, Columbia University, New York

\end{thebibliography}